---
title: "fynbos_glmer"
author: "Christopher Shortland"
date: '2022-06-11'
output: rmarkdown::github_document
---

# Libraries
```{r}
library(tidyverse)
library(lme4)
library(car)
library(effects)
library(ggplot2)
library(beepr)
library(sjPlot)
library(glmmTMB)
```
#importing final dataset
``` {r}
final <- read_csv('output/final_df/final_df.csv')
```
# Load fynbos stacked df
``` {r}
fyn_df <- filter(final, indicator=='fynbos')
```
#number of indicator species
``` {r}

fyn_df %>% filter(!Common_name %in% c('Jaeger, Pomarine',"Gull, Sabine's", "Chaffinch, Common", "Stint, Temminck's", "Scrub Robin, Rufous-tailed", "Partridge, Chukar", "Gull, Hartlaub's", "Cormorant, Crowned")) -> fyn_df

length(unique(fyn_df$Common_name))
unique(fyn_df$Common_name)
```
# add in year variable
``` {r}
fyn_df$Year <- lubridate::year(fyn_df$StartDate)
str(fyn_df)
```
# Option to run with Indicator Type as a fixed effect
# glmer1 <- glmer(Presence ~ Year*Indicator_Type + (1 + Year|Species_Code), family = 'binomial', data = fyn_df)

# measure time taken...
``` {r}
start_time <- Sys.time()
```
# run a mixed effects model; year = fixed; species = random
``` {r}
glmer1 <- glmer(Presence ~ Year + (Year|Species_Code) + (1|Pentad), family = 'binomial', data = fyn_df)
beepr::beep()
end_time <- Sys.time()
```
# How long did this take to run?
``` {r}
end_time - start_time
```
# Summary of model
``` {r}
summary(glmer1)
```
# Anova
``` {r}
Anova(glmer1)
```
# Plot effects
``` {r}
plot(allEffects(glmer1))
```
#predict
``` {r}
fyn_df$predict <- predict(glmer1, fyn_df)

fyn_clean <- fyn_df %>% group_by(Species_Code, Year) %>% summarise(predict = mean(predict))

ggplot(data=fyn_clean, aes(Year, predict)) + geom_line(aes(color = as.factor(Species_Code)))
```
``` {r}
plot_model(glmer1, type = "re")
```