---
title: "forest_glmer"
author: "Christopher Shortland"
date: '2022-05-24'
output: rmarkdown::github_document
---

# Libraries
```{r}
library(tidyverse)
library(lme4)
library(car)
library(effects)
library(ggplot2)
library(beepr)
```
#importing final dataset
``` {r}
final <- read_csv('output/final_df/final_df.csv')
```
# Load fynbos stacked df
```{r}
forest_df <- filter(final, indicator=='forest')
```
#number of indicator species
```{r}
length(unique(forest_df$Species_Code))
```
# add in year variable
```{r}
forest_df$Year <- lubridate::year(forest_df$StartDate)
str(forest_df)
```
# Option to run with Indicator Type as a fixed effect
## glmer1 <- glmer(Presence ~ Year*Indicator_Type + (1 + Year|Species_Code), family = 'binomial', data = forest_df)

# measure time taken...
```{r}
start_time <- Sys.time()
```
# run a mixed effects model; year = fixed; species = random
```{r}
forest_glmer <- glmer(Presence ~ scale(Year)  + (scale(Year)|Species_Code) + (1|Pentad), family = 'binomial', data = forest_df)
beepr::beep()
end_time <- Sys.time()
```
# How long did this take to run?
```{r}
end_time - start_time
```
# Summary of model
```{r}
summary(forest_glmer)
```
# Anova
```{r}
Anova(forest_glmer)
```
# Plot effects
```{r}
plot(allEffects(forest_glmer))
```
#predict
```{r}
forest_df$predict <- predict(forest_glmer, forest_df)

forest_clean <- forest_df %>% group_by(Species_Code, Year) %>% summarise(predict = mean(predict))
```
```{r}
ggplot(data=forest_clean, aes(Year, predict)) + geom_line(aes(color = as.factor(Species_Code)))
```
