---
title: "Karoo GLMER"
author: "Christopher Shortland"
date: '2022-05-24'
output: rmarkdown::github_document
---

# Libraries

``` {r}
    library(tidyverse)
    library(lme4)
    library(car)
    library(effects)
    library(ggplot2)
    library(beepr)
```

# Importing final dataset
``` {r}
        final <- read_csv('output/final_df/final_df.csv')
```
# Load fynbos stacked df
``` {r}
        karoo_df <- filter(final, indicator=='karoo')
```
#number of indicator species

``` {r}
    length(unique(karoo_df$Species_Code))
```

# add in year variable
``` {r}
        karoo_df$Year <- lubridate::year(karoo_df$StartDate)
        str(karoo_df)
```
# Option to run with Indicator Type as a fixed effect
glmer1 <- glmer(Presence ~ Year*Indicator_Type + (1 + Year|Species_Code), family = 'binomial', data = karoo_df)

# measure time taken...
``` {r}
        start_time <- Sys.time()
```
# run a mixed effects model; year = fixed; species = random

``` {r}
    glm_karoo <- glmer(Presence ~ scale(Year)  + (scale(Year)|Species_Code) + (1|Pentad), family = 'binomial', data = karoo_df)
    beepr::beep()
    end_time <- Sys.time()
```

# How long did this take to run?

``` {r}
    end_time - start_time
```

# Summary of model

``` {r}
    summary(glm_karoo)
```

# Anova

``` {r}
    Anova(glm_karoo)
```

# Plot effects

``` {r}
    plot(allEffects(glm_karoo))
```

# Predict
``` {r}
karoo_df$predict <- predict(glm_karoo, karoo_df)

karoo_clean <- karoo_df %>% group_by(Species_Code, Year) %>% summarise(predict = mean(predict))
```
# Prediction plot

``` {r}
    ggplot(data=karoo_clean, aes(Year, predict)) + 
        geom_line(aes(color = as.factor(Species_Code)))
```
